---
title: "Homework 4"
author: "Katie Miller"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Set Up

```{r download-packages}
# loading in the necessary packages
library(tidyverse)
library(here)
library(naniar)
library(dplyr)
library(performance)
library(broom)
library(flextable)
library(ggeffects)
library(car)
```

```{r read-in-data}
# reading in the data using here package
fish <- read_csv(here("data", "ntl6_v12.csv"))
```

```{r filter-data}
# filtering data frame to extract relevant info
trout <- fish %>% 
  filter(spname %in% "TROUTPERCH") %>% 
  reframe(year = case_when(spname == "TROUTPERCH" ~ year4),
          species = case_when(spname == "TROUTPERCH" ~ spname),
          length = case_when(spname == "TROUTPERCH" ~ length),
          weight = case_when(spname == "TROUTPERCH" ~ weight)) 

```

# Problem 1

## 1.

Biological:

Null: The length of a trout perch is not an accurate predictor of its weight.

Alternative: The length of a trout perch accurately predicts its weight.

Mathematical:

Null: the predictor variable does not predict the response variable.

Alternative: the predictor variable predicts the response variable.

## 2.
```{r visualize-missing-data}
# visualizing missing data in the filtered perch trout dataset
gg_miss_var(trout) + 
  labs(caption = "Plot 1: Missing data in the observations of perch trout. There are 199 missing\nmeasurements for trout perch weight, of the 489 observations, this is about 40%.\nThis missingness in our predicted variable lowers the sample size of our dataset.")
```
 

 
 ## 3.
```{r}
# creating a linear model for the trout perch dataset 
trout_model <- lm(weight ~ length, data = trout)
```

## 4.

```{r}
# plotting the trout model in a 2 by 2 grid
par(mfrow = c(2, 2))
plot(trout_model)
```

```{r}
dev.off()
```

```{r}
check_normality(trout_model)
check_heteroscedasticity(trout_model)
```


## 5. 

Residuals vs Fitted: the residuals are somewhat evenly distributed. There is a relatively even distribution above and below the line, however, the points are concentrated towards the middle of the plot. 

Scale-Location: there is slightly more homoscedasticity of variance in this plot, however there is still a bias towards the middle. 

Normal Q-Q: the residuals are fairly normally distributed. The points follow a linear path until a slight deviation around positive 2. 

Residuals vs Leverage: there are a few outliers measured influential by Cook's distance.

## 6.
```{r}
# creating a summary of the trout perch model
model_summary <- summary(trout_model)

model_summary
```

## 7.
```{r}
# setting up the initial ANOVA table
model_anova <- anova(trout_model)
```

```{r}
# creating a summary table
model_anova_table <- tidy(model_anova) %>% 
  mutate(across(sumsq:meansq, ~ round(.x, digits = 2))) %>% 
  mutate(statistic = round(statistic, digits = 1)) %>% 
  mutate(p.value = case_when(p.value < 0.001 ~ "< 0.001")) %>% 
   mutate(term = case_when(term == "length" ~ "Trout Perch Length", 
                           TRUE ~ term)) %>%
  flextable() %>% 
  set_header_labels(df = "Degrees of Freedom", 
                    sumsq = "Sum of squares",
                    meansq = "Mean squares",
                    statistic = "F-statistic",
                    p.value = "p-value")
  
model_anova_table
```

## 8.

connection between ANOVA table and summary object

## 9.
context (i.e. what do these stats mean in
the context of the original question)?
null hypothesis
test
sample size
test statistic
degrees of freedom
p-value
significance level
R2
interpretation of model equation

## 10.
```{r}
# creating the initial predictions object
predictions <- ggpredict(trout_model, terms = "length")
```

```{r}
predict_plot <- ggplot(data = trout,
                       aes(x = length, y = weight)) +
  geom_point() +
  geom_line(data = predictions, 
            aes(x = x, y = predicted), 
            color = "purple", linewidth = 1) +
  geom_ribbon(data = predictions, 
              aes(x = x, y = predicted, ymin = conf.low, 
                  ymax = conf.high), alpha = 0.3) +
  theme_bw() +
  labs(x = "Trout perch length", y = "Trout perch mass")


predict_plot
```


